---
title: "InterpretationTlessC"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Imprecision Interpretation TlessC preliminary analysis

Loading preprocessed data and extract critical trials

```{r tlessc}
df <- read.csv("Jan-26-2017-Batch_2666254_batch_results_intp_preprocessed.csv")
dfcrit <- df[df$SceneID %in% 1:40, ]

# drop levels in filler conditions
dfcrit$Condition <- factor(dfcrit$Condition)
dfcrit$Answer.choicePrior <- factor(dfcrit$Answer.choicePrior)
dfcrit$Answer.choiceAdj <- factor(dfcrit$Answer.choiceAdj)
dfcrit$Answer.choiceWhole <- factor(dfcrit$Answer.choiceWhole)
dfcrit$AdjectiveType <- factor(dfcrit$AdjectiveType)

```

## Choices after adjectives


```{r}
xtabs(~Condition + Answer.choiceAdj + AdjectiveType, dfcrit)
```

We can see that in general the presence of contrast boosts the chance of choosing the target, espeically
 for relative adjectives. 
 
## Choices after whole sentences
 

```{r}
dfwrong <- dfcrit[dfcrit$Answer.choiceWhole != "Target", ]
sort(table(as.factor(dfwrong$SceneID)), decreasing = TRUE)
``` 

We can see that 86 out of 100 participants made the "wrong" choice for Scene 86 even after seeing the whole sentence! 
And 74 got it wrong for Scene 13. What happened?!

Using the item visualizer (or just checking the stimuli lists), we find that in both cases, the participants chose the perfectly
 straight arrow when they were told to click on the straight line. 
In fact, the top 8 scenes in this list are all include lines and arrows and many participants just did not seem to make a distinction.

```{r}
sort(table(dfwrong$WorkerId), decreasing = TRUE)
``` 
There is at least a small group of participants who do not distinguish between
lines and arrows. 
I do not have a good sense of how likely an English speaker will make such a distinction, but I think it could be 
that the participants "learned" to ignore such a distinction to accomodate the speaker. (One participant commented "You used arrow and line interchangeably and I found that confusing.")

Note that participants seemed more willing to give up this distinction when the adjective is "straight" then when it is a relative adjective.
However, it is unclear what the overall effect is, since there are more relative adjectives scenes using lines than absolute adjectives.

```{r}
itemResponses <- function(df, id){
  r = rbind(table(df$Answer.choicePrior[df$SceneID==id]), 
        table(df$Answer.choiceAdj[df$SceneID==id]), 
        table(df$Answer.choiceWhole[df$SceneID==id]))
  rownames(r) = c('prior', 'adj', 'noun')
  return(r)
}
```